CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-12-04 codecheck-paper.tex
extend limitations

\textbf{Just peer review}: Wait, isn't that what peer review should be 
doing already? That is correct.
The general observation is that the increasing
number of researchers, submissions, and reviews to conduct under the
pressure to publish, puts the scholarly publication system under a lot
of strain [REF].  We also see parallels to the terms 
`reproducible research' and `open science'. Establishing a CODECHECK 
process is largely an acknowledgement that peer reviewing practices 
have faltered to adopt to the challenges of digitisation and computer-based
research or data science. The concept of a CODECHECK, just as the 
reproducibility of research and the openness of science, may be of 
transitional nature. If the activites described here as being part of a 
CODECHECK become the norm for scientific reviews and publications, then
the initiative will have succeeded.
However, the rapidly changing technology used in  research leads us to 
believe that a seperate role of codechecker is benefitial in general.

\textbf{Handling failure}: we have no established process for this, as
so far all our codechecks succeeded. Open question remains as what to do
in this case -- should we publicly report that the code would not run.
Be careful as to the impact of acceptance or volunteers this decision
has.

@@

\textbf{Software}: authors can currently provide code that requires
propieratiry software. Given the prevalence of software like MATLAB,
this seems like a pragmatic choice, but then (a) we can't rely on open
infrastructure like mybinder, (b) codechecker must have that software.

\textbf{Can't someone cheat?} Yes. we don't check it is correct code,
just simply that it runs. This `mechanical' test is indeed a very low bar.
Yet, by having the code and data openly deposited, interested parties can
then examine the code, and knowing that the code will be open we hope will
be an incentive for authors to share. It will also give the opportunity 
to future researchers, potentially with new methods, to look for errors.
This is a much more effective means to avoid cheating than an arms race
trying to catch the cheater with closed datasets and code.
% give credit to Everything Hertz podcast for the idea? they compared it
% to the blood samples of drug tests in sports (stored for n years)

\textbf{Who's got time for more peer review?} Agree, but ECRS might be
attracted to working with groups (recall it is not anonymous)

\textbf{Who finds the codechecker?} Good point! The journal. Community
however think this is an important issue, and are volunteering.

\textbf{Checking multiple times?}  If you check a preprint before peer
review, it might need checking again as it is modified during peer
review.  Unavoidable, and as happened to us (2020-012 has been
revised).  But checking the manuscript the second time is likely to be
much less work than the first time.

>>>>

\textbf{Just peer review}: wait, isn't CODECHECK what peer review should be 
doing already? That is correct.
However, the general observation is that the increasing
number of researchers, submissions, and reviews to conduct under the
pressure to publish, puts the scholarly publication system under a lot
of strain [REF].  We also see parallels to the terms 
`reproducible research' and `open science'. Establishing a CODECHECK 
process is largely an acknowledgement that peer reviewing practices 
have faltered to adapt to the challenges of digitisation and computer-based
research or data science. The concept of a CODECHECK, just as the 
reproducibility of research and the openness of science, may be of 
transitional nature. If the activites described here as being part of a 
CODECHECK become the norm for scientific reviews and publications, then
the initiative will have succeeded.
However, the rapidly changing technology used in research leads us to 
believe that a seperate role of codechecker and the expertise it brings
will be benefitial in the long term.

\textbf{Not going deep enough}: 
It should be noted that a codechecker is not taking the same roles as the
\emph{statistical reviewer}, as it is practices by some journals in the 
biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
The statistical reviewer actually evaluates the appropriateness of
statistical methods \cite{greenwood_how_2015} and can support topical
reviewers if, e.g., complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.
The codechecker may go equally deep into the review, if they have, by 
intention of by chance, the required expertise and the required time, 
but they are by no means expected to be able to provide
such a level of scrutiny. We can imagine a tiered workflow where the 
codechecker, just as the conventional reviewer, could recommend a detailed
statistical review to the editor as they might come across upon different
hints.
%The CODECHECK provides a certain level of safety with regard to the inherent
%problems that a computational workflow might have. We want to capture the
%obvious errors, but ... https://www.quora.com/What-is-the-difference-between-safety-and-security

\textbf{Handling failure}: we have no established process for the case
that a reproduction fails. This may be the sign of a pre-selection bias
of the examples, as so far all our codechecks succeeded.
In case of a journal adopting CODECHECK for all submissions, the open 
question remains as what to do in this case will have to be answered.
At the start, we doubt publicly reporting failures (i.e., the code 
would not run) will increase overall reproducibility and recommend not
to share the outcome publicly but to share the latest draft of the 
certific with the authors.
One must be careful as to the impact of acceptance by authors and the
concerns of volunteers this approach, or a contrary one, has.

@@

\textbf{Proprietary software}: authors can currently provide code that requires
proprietary software. Given the prevalence of proprietary software in
some disciplines, such as MATLAB or eCognition,
this seems like a pragmatic choice.
However, it prohibits to benefit from open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020}) and requires the codechecker to have access 
to that software.
Non-open software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a compromise
that each implementation of CODECHECK should critically reconsider.

\textbf{Can't someone cheat?} Yes. We don't check it is correct code,
just simply that it runs. This `mechanical' test is indeed a very low bar.
Yet, by having the code and (raw) data openly deposited, interested parties can
then examine the code, and knowing that the code will be open we hope will
be an incentive for authors to share. It will also give the opportunity 
to future researchers, potentially with new methods, to look for errors.
This is a much more effective means to avoid cheating than an arms race
trying to catch the cheater with closed datasets and code. This approach
is comparable to the storing of blood samples in sports to detect
doping in the future (cf. \cite{everythinghertz97}).
Another picture that helped us to define the scope of a CODECHECK is:
the codechecker is the forensic photographer, making sure all the details
are captured so that an investigator may, now or in the future, scrutinize
them.

\textbf{Who's got time for more peer review?} Agree, but the technical
nature and specific skill set might allow to introduce different groups
into the peer review process. ECRs might be attracted to working with
groups to learn more about recent methods and reproducibility practices.
Research software experts (Research Software Engineers, RSE) who might
not regularly be involved in writing or reviewing papers could be 
interested to increase their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people, who must not even
be traditional peer reviewers. In fact, the codechecker could even be
a software professional employed by a publisher.
Also, recall that we expect CODECHECKSs to be not anonymous and always
publicly deposited, which is common practice for some journals, but not
broadly practiced. We also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very educational. It's a different role and as such could
be a welcome alternation/diversion for reviewers.

\textbf{Who finds the codechecker?} Good point! The journal, using either
their own team of codecheckers or the public database. Community
however think this is an important issue, and are volunteering.

\textbf{Checking multiple times?} If you check a preprint before peer
review or a submission to a journal with non-public review,
it might need checking again as it is modified during peer
review.  This is in part unavoidable, and as happened to us (e.g., 
`2020-012` has been revised). This can also be desired, if the
interaction between author, reviewer, and codechecker lead to
improvements.
Since checking the manuscript the second time is likely to be
much less work than the first time, this is largely mitigated by 
open and regular communication.
